---
title: Working With Data
date: today
date-format: long
footer: "[DataViz 2102 Website](https://dataviz-gwu.rocks)"
logo: images/dataviz-logo.png
format:
  revealjs:
    theme: [simple, custom.scss]
    transition: fade
    slide-number: true
    #multiplex: true
    chalkboard: true
execute:
  echo: true
  message: false
  warning: false
  freeze: auto
---

## How Do We Get Tidy/Clean Data? 

<br>

- Wrangle it ourselves
- Use a package where it has been wrangled for us
- Download via an API 

## This Lesson

<br>

- Practice with World Bank and V-Dem data
- World Bank data through `wbstats`
  - There is another package called `WDI`
  - Both packages for accessing data through WB API
- Varieties of Democracy (V-Dem) through `vdemdata`
  - Downloads the full V-Dem dataset, which we pare down with `dplyr`
  
## `filter()`, `select()`, `mutate()` {.smaller}

<br>

Along the way we will practice some important `dplyr` verbs:

<br>

- `filter()` is used to select observations based on their values
- `select()` is used to select variables
- `mutate()` is used to create new variables or modifying existing ones  

<br>

As well as some helpful functions from the `janitor` package.  

## APIs

<br>

- API stands for "Application Programming Interface"
- Way for two computers to talk to each other
- In our case, we will use APIs to download social science data

## APIs in R

::: {.incremental}
- APIs are accessed through packages in R
- Sometimes there can be more than one package for an API
- Much easier than reading in data from messy flat file!
- We will use a few API packages in this course
  - World Bank data through `wbstats` (or `WDI`)
  - `fredr` for Federal Reserve Economic Data
  - `tidycensus` for US Census data
- But there are many APIs out there (please explore!)
:::


## Searching for WB Indicators

<br>

```{r}
#| label: search_wb
#| eval: false

flfp_indicators <- wb_search("female labor force") # store the list of indicators

print(flfp_indicators, n=26) # view the indicators
```

## `wbstats` Example

<br>

```{r}
#| label: download_wb
#| eval: false
#| echo: true
#| code-line-numbers: "1-4|6-7|9|10|11|12|13-16|18-19"

# Load packages
library(wbstats) # for downloading WB data
library(dplyr) # for selecting, renaming and mutating
library(janitor) # for rounding

# Store the list of indicators in an object
indicators <- c("flfp" = "SL.TLF.CACT.FE.ZS", "women_rep" = "SG.GEN.PARL.ZS") 

# Download the data  
women_emp <- wb_data(indicators, mrv = 50) |> # download data for last 50 yrs
  select(!iso2c) |> # drop the iso2c code which we won't be using
  rename(year = date) |> # rename date to year 
  mutate(
    flfp = round_to_fraction(flfp, denominator = 100), # round to nearest 100th
    women_rep = round_to_fraction(women_rep, denominator = 100) 
  )

# View the data
glimpse(women_emp) 
```

## Your Turn!

<br>

- Search for a WB indicator
- Download the data

```{r}
#| label: timer1
#| echo: false

library(countdown)
countdown(minutes = 5, 
          id = "timer1", 
          top = "10%", 
          right = "10%",
          color_border = "#fff",
          color_text = "#fff",
          color_running_background = "#42affa",
          color_running_text = "black",
          color_finished_background = "#E5D19D",
          color_finished_text = "#00264A")
```

## The V-Dem Dataset

<br>

- V-Dem stands for Varieties of Democracy
- It is a dataset that measures democracy around the world
- Based on expert assessments of the quality of democracy in each country
- We will access V-Dem data through the `vdemdata` package

## Downloading V-Dem Data

<br>

The `vdem` function from `vdemdata` just downloads all of the data. Try running this code chunk. What do you see in `democracy`?

```{r}
#| label: vdemdata1
#| eval: false

library(vdemdata) # load the V-Dem package

democracy <- vdem() # download the V-Dem dataset
```


## `filter()`

- Run this code. What do you see?
- Try changing the year 
- For one year, use `==` instead of `>=`
- Or try `<=` and see what happens 

```{r}
#| label: vdemdata2
#| eval: false

democracy <- vdem |> # download the V-Dem dataset
  filter(year >= 1990) # filter out years less than 1990
  
glimpse(democracy)  
```

## `=` versus `==`

<br>

- `=` is used to assign values to variables, just like `<-`
- `==` is used to test if two values are equal to each other
- So `filter(year == 1990)` will give you just the observations for 1990

## `>=` and `<=`

- `>=` is used to test if a value is greater than or equal to another value
- `<=` is used to test if a value is less than or equal to another value
- So `filter(year >= 1990)` will give you the observations for 1990 and later
- And `filter(year <= 1990)` will give you the observations for 1990 and earlier

## `select()`

- Run this code. What do you see?
- Now try `v2x_libdem` instead of `v2x_polyarchy`
- Choose more from the [codebook](https://v-dem.net/documents/24/codebook_v13.pdf) 

```{r}
#| label: vdemdata3
#| echo: true
#| eval: false

democracy <- vdem |> # download the V-Dem dataset
  select(                  # select (and rename) these variables
    country = country_name,     # before the = sign is new name  
    vdem_ctry_id = country_id,  # after the = sign is the old name
    year, 
    polyarchy = v2x_polyarchy
  )
  
glimpse(democracy)  
```

```{r}
#| label: timer2
#| echo: false

countdown(minutes = 5, 
          id = "timer2",
          top = "0%",
          right = "5%")
```

## `mutate()`

- Modify the code to create new variable that is three times the value of `polyarchy` 
- How about `polyarchy` squared? 

```{r}
#| label: vdemdata4
#| eval: false

democracy <- vdem |> # download the V-Dem dataset
  filter(year == 2015) |> # keep only observations from 2015
  select(                  # select (and rename) these variables
    country = country_name,     # name before the = sign is new name  
    vdem_ctry_id = country_id,  # name after the = sign is old name
    year, 
    polyarchy = v2x_polyarchy 
    ) |>
  mutate(
    polyarchy_dbl = polyarchy * 2 # create variable 2X polyarchy
  )
  
glimpse(democracy)  
```

```{r}
#| label: timer3
#| echo: false

countdown(minutes = 5, 
          id = "timer3",
          top = "0%",
          right = "5%")
```

## Some Common Arithmetic Operators

<br>

- `+` addition
- `-` subtraction
- `*` multiplication
- `/` division
- `^` exponentiation (also `**`)

## `vdemdata` Example {.smaller}

<br>

```{r}
#| label: vdemdata5
#| eval: false

# Load packages
library(vdemdata) # to download V-Dem data
library(dplyr)

# Download the data
democracy <- vdem |> # download the V-Dem dataset
  filter(year == 2015)  |> # filter year, keep 2015
  select(                  # select (and rename) these variables
    country = country_name,     # the name before the = sign is the new name  
    vdem_ctry_id = country_id,  # the name after the = sign is the old name
    year, 
    polyarchy = v2x_polyarchy, 
    gdp_pc = e_gdppc, 
    region = e_regionpol_6C
    ) |>
  mutate(
    region = case_match(region, # replace the values in region with country names
                     1 ~ "Eastern Europe", 
                     2 ~ "Latin America",  
                     3 ~ "Middle East",   
                     4 ~ "Africa", 
                     5 ~ "The West", 
                     6 ~ "Asia")
  )

# View the data
glimpse(democracy)
```

## {.smaller}

<br>

Use `filter()` to select years...

<br>

```{r}
#| label: vdemdata6
#| eval: false
#| code-line-numbers: "2-3"

# Download the data
democracy <- vdem |> 
  filter(year == 2015)  |> # keep 2015
  select(                 
    country = country_name,       
    vdem_ctry_id = country_id,  
    year, 
    polyarchy = v2x_polyarchy, 
    gdp_pc = e_gdppc, 
    region = e_regionpol_6C
    ) |>
  mutate(
    region = case_match(region,
                     1 ~ "Eastern Europe", 
                     2 ~ "Latin America",  
                     3 ~ "Middle East",   
                     4 ~ "Africa", 
                     5 ~ "The West", 
                     6 ~ "Asia")
  )
```

## {.smaller}

<br>

Use `select()` to choose variables...

<br>

```{r}
#| label: vdemdata7
#| eval: false
#| code-line-numbers: "4-11"

# Download the data
democracy <- vdem |> 
  filter(year == 2015)  |> 
  select(                  # select (and rename) these variables
    country = country_name,     # the name before the = sign is the new name  
    vdem_ctry_id = country_id,  # the name after the = sign is the old name
    year, 
    polyarchy = v2x_polyarchy, 
    gdp_pc = e_gdppc, 
    region = e_regionpol_6C
    ) |>
  mutate(
    region = case_match(region, 
                     1 ~ "Eastern Europe", 
                     2 ~ "Latin America",  
                     3 ~ "Middle East",   
                     4 ~ "Africa", 
                     5 ~ "The West", 
                     6 ~ "Asia")
  )
```

## {.smaller}

<br>

Use `mutate` with `case_match()` to Recode Region....

<br>

```{r}
#| label: vdemdata8
#| eval: false
#| code-line-numbers: "12-23"

# Download the data
democracy <- vdem |>
  filter(year == 2015)  |> 
  select(                  
    country = country_name,     
    vdem_ctry_id = country_id,  
    year, 
    polyarchy = v2x_polyarchy, 
    gdp_pc = e_gdppc, 
    region = e_regionpol_6C
    ) |>
  mutate(
    region = case_match(region, # replace the values in region with country names
                     1 ~ "Eastern Europe", 
                     2 ~ "Latin America",  
                     3 ~ "Middle East",   
                     4 ~ "Africa", 
                     5 ~ "The West", 
                     6 ~ "Asia")
                    # number on the left of the ~ is the V-Dem region code
                    # we are changing the number to the country name on the right
                    # of the equals sign
  )
```

## Your Turn

<br>

- Use `vdem` to download the full V-Dem dataset
- Use `filter()` to keep a year (or years) you are interested in
- Use `select()` to choose `country_name`, `year`, and an indicator of interest
  - e.g., `v2x_polyarchy` for electoral democracy
- Use `mutate()` with `case_match()` to recode the region variable (`e_regionpol_6C`)
- `glimpse()` your data to check the result


```{r}
#| label: timer4
#| echo: false

countdown(minutes = 5, 
          id = "timer4",
          top = "0%",
          right = "5%")
```