---
title: "Themes, Annotations and Interactivity"
date: today
date-format: long
footer: "[DataViz 2102 Website](https://dataviz-gwu.rocks)"
logo: images/dataviz-logo.png
format:
  revealjs:
    theme: [simple, custom.scss]
    transition: fade
    slide-number: true
    #multiplex: true
    chalkboard: true
execute:
  echo: true
  message: false
  warning: false
  freeze: auto
---

```{r}
#| label: setup
#| include: false

library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(wbstats)
library(countrycode)
library(countdown)

indicators = c(flfp = "SL.TLF.CACT.FE.ZS", gdp_pc = "NY.GDP.PCAP.KD") # define indicators

## Line chart from last class

dem_waves_ctrs <- read_csv("data/dem_waves_ctrs.csv")

## Cross-section of data on FLFP and GDP per capita for scatter plot

flfp_gdp <- wb_data(indicators, end_date = 2021) |> # download data for 2021
    left_join(select(wb_countries(), c(iso3c, region)), by = "iso3c") |>  # add regions
    drop_na() # drop rows with missing values

## Time series data on regional trends in FLFP for line chart

flfp_ts <- wb_data("SL.TLF.CACT.FE.ZS", country = "regions_only", start_date = 1990, end_date = 2022) |> 
  rename(
    region = country,
    year = date,
    flfp = SL.TLF.CACT.FE.ZS
  ) |> 
  select(region, iso3c, year, flfp)
```

## Line Chart from Last Class

```{r}
#| label: line_chart1
#| echo: false

dem_waves_chart <- ggplot(dem_waves_ctrs, aes(x = year, y = polyarchy, color = country)) +
  geom_line(linewidth = 1) +
  labs(
    x = "Year",
    y = "Polyarchy Score",
    title = 'Democracy in countries representing three different "waves"',
    caption = "Source: V-Dem Institute",
    color = "Country"
    ) +
  scale_color_viridis_d(end = .8)

dem_waves_chart
```

## Add a Theme

```{r}
#| label: line_chart2
#| code-fold: true

dem_waves_chart + theme_minimal()
```

## Can Also Set Theme {.smaller}

<br>

If you like particular theme and want to use it for all of your visualizations, use `theme_set()` in a code chunk at the top of your document, e.g....

<br>

```{r}
#| label: my_setup_chunk
#| echo: fenced

theme_set(theme_minimal())
```

<br>

... and R will apply that theme to all of the visualizations in your document. 

## Your Turn! 

Produce this scatter plot... Then try different [themes](https://ggplot2.tidyverse.org/reference/ggtheme.html).

```{r}
#| label: exercise1
#| echo: false

wealth_flfp <- ggplot(flfp_gdp, aes(x = gdp_pc, y = flfp)) + 
  geom_point(aes(color = region)) + # color points by region
  geom_smooth(method = "loess", linewidth = 1) +  # make the line a loess curve
  scale_x_log10(labels = scales::label_dollar()) + # stretch axis, add '$' format
  scale_y_continuous(labels = scales::label_percent(scale = 1)) + # add % label
  labs(
    x= "GDP per Capita", # x-axis title
    y = "Female Labor Force Participation", # y-axis title
    title = "Wealth and female labor force participation, 2021", # plot title
    caption = "Source: World Bank Development Indicators", # caption
    color = "Region" # legend title
    ) + 
  scale_color_viridis_d(option = "plasma", end = .7) 

wealth_flfp + theme_dark()
```

```{r}
#| label: setup_timer
#| echo: false

countdown(minutes = 5, 
          id = "setup", 
          top = "10%", 
          right = "-15%",
          color_border = "#fff",
          color_text = "#fff",
          color_running_background = "#42affa",
          color_running_text = "black",
          color_finished_background = "#E5D19D",
          color_finished_text = "#00264A")
```

## Add an Annotation

```{r}
#| label: annotation1
#| echo: false

flfp_line <- ggplot(flfp_ts, aes(x = year, y = flfp, color = region)) +
  geom_line(linewidth = 1) + 
  scale_y_continuous(labels = scales::label_percent(scale = 1)) +
  labs(
    x = "Year", 
    y = "Female Labor Force Participation", 
    title = "Regional trends in female labor force participation", 
    caption = "Source: World Bank", 
    color = "Region"
  ) + scale_color_viridis_d() +
  theme_minimal()

flfp_line + geom_hline(yintercept = 52, linetype = "dashed", color = "red", linewidth = 1) +
  annotate("text", x = 1995, y = 55, label = "Global average")
```

## 

First step, add a horizontal line using `geom_hline()`... 

```{r}
#| code-fold: true

flfp_line + 
  geom_hline(yintercept = 52, linetype = "dashed", color = "red", linewidth = 1) 
```

## 

Second step, the annotation with `annotate()`...

```{r}
#| code-fold: true 

flfp_line + 
  geom_hline(yintercept = 52, linetype = "dashed", color = "red", linewidth = 1) +
  annotate("text", x = 1995, y = 55, label = "Global average")
```

## Vertical Reference Line

```{r}
#| echo: false

flfp_line + 
  geom_vline(xintercept=2020, linetype = "dashed", linewidth = 1) +
  annotate("text", x = 2017, y = 35, label = "Pandemic")
```

## 

First step, add a horizontal line using `geom_hline()`... 

```{r}
#| code-fold: true

flfp_line + 
  geom_vline(xintercept=2020, linetype = "dashed", linewidth = 1)
```

## 

Second step, add the annotation with `annotate()`

```{r}
#| code-fold: true

flfp_line + 
  geom_vline(xintercept=2020, linetype = "dashed", linewidth = 1) +
  annotate("text", x = 2017, y = 35, label = "Pandemic")
```

## Your Turn!

<br>

- Go to module 2.2
- Copy the code you need to make the line chart
- Create a horizontal reference line and annotate it
- In a new code chunk, create a vertical line and annotate it
- Play with the parameters to move and style your line
- Try different text, etc. 

```{r}
#| label: exercise2_timer
#| echo: false

countdown(minutes = 5, 
          id = "exercise2", 
          #bottom = "10%", 
          #right = "10%"
          )
```

## Interactivity

<br>

Any ggplot can become an interactive chart with one function: `ggplotly()` from the `plotly` package.

```{r}
#| label: plotly_setup
#| echo: false

library(plotly)

wealth_flfp_plotly <- wealth_flfp +
  scale_color_viridis_d(option = "plasma", end = .7) +
  theme_minimal()
```

```{r}
#| label: plotly_basic
#| echo: true
#| output: false

library(plotly)

ggplotly(wealth_flfp_plotly)
```

##

```{r}
#| label: plotly_basic_out
#| echo: false

ggplotly(wealth_flfp_plotly)
```

## Controlling the Tooltip {.smaller}

<br>

By default plotly shows every mapped variable. Use `tooltip =` to pick which ones, and add `aes(label = )` for a variable that is not already in the plot...

```{r}
#| label: plotly_tooltip
#| echo: true
#| output: false

wealth_flfp_plotly <- wealth_flfp_plotly + aes(label = country) # so plotly keeps country

ggplotly(wealth_flfp_plotly, tooltip = c("country", "gdp_pc", "flfp"))
```

##

```{r}
#| label: plotly_tooltip_out
#| echo: false

ggplotly(wealth_flfp_plotly + aes(label = country),
         tooltip = c("country", "gdp_pc", "flfp"))
```

## Your Turn!

- Go to module 2.2 and create the basic wealth vs. flfp scatter plot
- Now "plotlify" it by wrapping it in `ggplotly()`
- Hover over the points to see the default tooltip
- Now use `tooltip =` to choose which variables appear
- Try adding `aes(label = country)` so you can show the country name

```{r}
#| label: exercise3_timer
#| echo: false

countdown(minutes = 5, 
          id = "exercise3", 
          #bottom = "10%", 
          #right = "10%"
          )
```

## Animated Plots

```{r}
#| label: animate_out
#| echo: false

flfp_gdp_ts <- wb_data(indicators, mrv = 25) |> # download data for 25 most recent years
    left_join(select(wb_countries(), c(iso3c, region)), by = "iso3c") |> # add regions
    rename(year = date) |>
    drop_na() # drop rows with missing values

library(gganimate)

ggplot(flfp_gdp_ts, aes(x = gdp_pc, y = flfp)) +
  geom_point(aes(color = region)) +
  geom_smooth(method = "loess", linewidth = 1) +
  scale_x_log10(labels = scales::label_dollar()) +
  scale_y_continuous(labels = scales::label_percent(scale = 1)) +
  transition_states(year) +
  labs(
    x= "GDP per Capita",
    y = "Female Labor Force Participation",
    title = "Wealth and female labor force participation, {closest_state}",
    caption = "Source: World Bank Development Indicators",
    color = "Region"
    ) +
  scale_color_viridis_d(option = "plasma", end = .7)
```

## Animated Plots {.smaller}

Two additions: `transition_states()` picks the variable to animate over, and `{closest_state}` puts it in the title.

```{r}
#| label: animate_code
#| eval: false
#| echo: true
#| code-line-numbers: "1|3-5|7|12"

library(gganimate)

flfp_gdp_ts <- wb_data(indicators, mrv = 25) |> # 25 most recent years
    left_join(select(wb_countries(), c(iso3c, region)), by = "iso3c") |>
    drop_na() |> rename(year = date)

ggplot(flfp_gdp_ts, aes(x = gdp_pc, y = flfp)) +
  geom_point(aes(color = region)) +
  scale_x_log10(labels = scales::label_dollar()) +
  scale_y_continuous(labels = scales::label_percent(scale = 1)) +
  transition_states(year) +
  labs(title = "Wealth and female labor force participation, {closest_state}") +
  scale_color_viridis_d(option = "plasma", end = .7)
```

## Your Turn!

<br>

- Try making an animated scatter plot with different a different y-axis variable
- See if you can make the title change based on the year displayed

```{r}
#| label: exercise4_timer
#| echo: false

countdown(minutes = 5, 
          id = "exercise4", 
          #bottom = "10%", 
          #right = "10%"
          )
```
