---
title: "Coding Assignment 3"
execute:
  echo: true
  message: false
  warning: false
---

## Overview

This assignment has two parts. In **Part I** you will use `tidycensus` and the `gt` package to build a polished table of census data. In **Part II** you will use `peacesciencer`, `broom`, and `modelsummary` to analyze civil-conflict onset and present the results as both a regression table and a coefficient plot.

## Part I: A census income table (40 pts)

*Choose a state you find interesting from the standpoint of income distribution. Using `tidycensus`, download county-level data on income quintiles (specify `geography = "county"` and your chosen `state = `). Clean the variable names and use `mutate(name = str_replace_all(...))` so the table shows just the county name rather than "X County, State."*

*Then use the `gt` package to make a polished table of the income quintiles by county: add a title and subtitle, relabel the columns, format the numbers as dollar figures, and add a source note. Beautify it as you see fit. Below the table, interpret it in a few sentences of markdown text---which counties stand out?*

```{r}

```

## Part II: Regression tables and coefficient plots (60 pts)

*For Part II we evaluate another classic: Fearon and Laitin's [Ethnicity, Insurgency, and Civil War](https://www.jstor.org/stable/3118222) (cited some 11,000 times). Their provocative thesis is that ethnic diversity per se is not an important predictor of civil conflict. We will approximate their analysis using `peacesciencer`.*

### Step 1: Build your data set (20 pts)

*Using `create_stateyears()` and the various "add" functions (`add_ucdp_acd()`, `add_democracy()`, and so on) as we did in class, assemble a state-year data frame for analyzing conflict onset, filtered to 1946--2019.*

```{r}

```

### Step 2: Run models and make a regression table (25 pts)

*Regress `ucdponset` on a set of predictors including a measure of ethnic fractionalization or polarization (`ethfrac`/`ethpol`), a democracy measure (`v2x_polyarchy` or `polity2`), terrain (`rugged` or `newlmtnest`), GDP per capita, and population. Fit two or three model specifications and present them side by side in a `modelsummary` table. How do your results compare to Fearon and Laitin's Table 1?*

```{r}

```

### Step 3: Coefficient plot (15 pts)

*Use `modelplot()` to display one of your models as point estimates with confidence intervals. In a sentence or two, discuss the tradeoffs of presenting results this way versus in a table.*

```{r}

```

## Submission

*After rendering your document, compress (zip) your project folder and submit it on Blackboard. You will find the submission portal under the Assignments link. There is a screen capture video in the Discord server that will help you understand how to do this.*
