---
title: "Coding Assignment 1"
execute:
  echo: true
  message: false
  warning: false
---

## Overview

In this assignment you will pull together everything from the first part of the course: you will **download** data from two sources, **wrangle and merge** it, and then **visualize** it to evaluate a classic question in comparative politics---the relationship between economic development and democracy famously argued by Seymour Martin Lipset in ["Some Social Requisites of Democracy."](https://www-jstor-org.proxygw.wrlc.org/stable/1951731?seq=8)

You will retrieve a measure of democracy from V-Dem and a measure of economic development, wrangle them into a single tidy data set, and then make a couple of charts to explore whether wealthier countries tend to be more democratic. Do your work in the code chunks provided, and write all interpretation as markdown text (not inside code chunks).

Here is a setup chunk. Load your packages here or as you go along.

```{r}
#| label: setup

```

:::{.callout-warning}
Please refrain from *installing* packages in your code chunks---installation should happen once in the console, not every time the document renders.
:::

## Step 1: Download and merge your data (30 pts)

*Choose one measure of democracy from V-Dem---for example liberal democracy (`v2x_libdem`), participatory democracy (`v2x_partipdem`), deliberative democracy (`v2x_delibdem`), or egalitarian democracy (`v2x_egaldem`)---and use `fetchdem()` from `vdemlite` to download it along with GDP per capita (`e_gdppc`) and region codes (`e_regionpol_6C`) for all countries from roughly 1990 onward.*

*If you would like an extra challenge, you may instead pull GDP per capita from the World Bank with `wbstats`, add country codes with the `countrycode` package, and merge it to your V-Dem data with `left_join()`. Either way, `glimpse()` the result to confirm everything came through.*

```{r}

```

## Step 2: Summarize your data (20 pts)

*Use a `group_by()`, `summarize()`, `arrange()` sequence to glean some insight from your data. For example, how do world regions compare on average democracy and average GDP per capita? Which countries stand out? Write a sentence or two below the chunk describing what you find.*

```{r}

```

## Step 3: Line chart of democracy over time (20 pts)

*Pick five countries from different regions and make a line chart of your democracy measure from 1990 onward using `ggplot2` and `geom_line()`. Map country to `color`, add clear axis labels, a title, and a caption, apply a colorblind-friendly palette and an attractive theme, and add one annotation (for example a vertical line marking an important date). Then interpret the chart in a few sentences of markdown text.*

```{r}

```

## Step 4: Scatter plot of development vs. democracy (20 pts)

*Using one year of your data, make a scatter plot with GDP per capita on the x-axis and your democracy measure on the y-axis, coloring the points by region. Put the x-axis on a log scale, use the `scales` package to format it as dollars, apply a `viridis` color map and a theme, and add a linear trend line with `geom_smooth(method = "lm")`. Add appropriate labels.*

```{r}

```

## Step 5: Interpretation (10 pts)

*Bring your charts together. Does the evidence support Lipset's claim that economic development and democracy go together? Note any interesting exceptions---countries that are wealthier than their level of democracy would predict, or vice versa---and say what they might tell us.*


## Extra Credit (3 pts)

*Make your scatter plot interactive with `ggplotly()`, making sure the tooltip shows the country name and the values you care about.*

```{r}

```

**Submission:** After rendering your document, compress (zip) your project folder and submit it on Blackboard. You will find the submission portal under the Assignments link. There is a screen capture video in the Discord server that will help you understand how to do this.
