---
title: "Lab 2: Wrangling"
execute:
  echo: true
  message: false
  warning: false
---

::: {.callout-tip}
## How labs work
Labs are done **with your group during class** and are meant to be mostly finished in the class period. You may use AI to **ask questions or debug**, but not to write your code for you. At the end, add a one-line note saying whether and how you used AI.
:::

## Overview

In Lab 1 we handed you clean data. Now you will build a tidy data set yourself---the skill that makes every later visualization possible. You will pull a democracy measure from V-Dem and a development or empowerment indicator from the World Bank, wrangle them, and merge them into one data frame.

```{r}
#| label: setup

```

## Step 1: Download your data (30 pts)

*Use `fetchdem()` from `vdemlite` to download a democracy measure of your choice for all countries from roughly 1990 onward. Then use `wb_data()` from `wbstats` to download GDP per capita (or another development or women's-empowerment indicator) for the same period. `glimpse()` each to confirm they came through.*

```{r}

```

## Step 2: Clean and merge (40 pts)

*Get both data frames into a compatible shape (rename or reshape as needed). Add country codes with the `countrycode` package, then join the two sources into a single data frame with `left_join()`. `glimpse()` the result and check for excessive `NA`s.*

```{r}

```

## Step 3: Summarize (30 pts)

*Use a `group_by()`, `summarize()`, `arrange()` sequence to pull one insight from your merged data---for example, how regions compare on average democracy and development, or which countries stand out. Write a sentence or two below the chunk describing what you find.*

```{r}

```


*AI note: _(one line on whether/how your group used AI)_*
