---
title: "Lab 4: Tables & coefficient plots"
execute:
  echo: true
  message: false
  warning: false
---

::: {.callout-tip}
## How labs work
Labs are done **with your group during class** and are meant to be mostly finished in the class period. You may use AI to **ask questions or debug**, but not to write your code for you. At the end, add a one-line note saying whether and how you used AI.
:::

## Overview

In this lab you will present data two ways: as a polished **table** with `gt`, and as a **regression table plus coefficient plot** with `modelsummary`. For Part II we approximate Fearon and Laitin's [Ethnicity, Insurgency, and Civil War](https://www.jstor.org/stable/3118222), whose provocative claim is that ethnic diversity per se does not predict civil conflict.

```{r}
#| label: setup

```

## Part I: A `gt` table (40 pts)

*Choose a state and use `tidycensus` to download county-level data on income quintiles (`geography = "county"`). Clean the variable names and use `mutate(name = str_replace_all(...))` so only the county name shows. Then make a polished `gt` table: add a title and subtitle, relabel the columns, format the numbers as dollars, and add a source note. Below the table, note in a sentence which counties stand out.*

```{r}

```

## Part II: Regression table and coefficient plot (60 pts)

*Use `create_stateyears()` and the `peacesciencer` "add" functions to build a state-year data frame for analyzing conflict onset (`ucdponset`), filtered to 1946--2019. Fit two or three model specifications---varying, say, ethnic fractionalization vs. polarization, the democracy measure, or the terrain measure---and present them side by side in a `modelsummary` table. Then use `modelplot()` to show one model as a coefficient plot with confidence intervals. In a sentence or two, compare your results to Fearon and Laitin's and note what the coefficient plot makes easier to see than the table.*

```{r}

```


*AI note: _(one line on whether/how your group used AI)_*
