---
title: "Lab 3: Maps"
execute:
  echo: true
  message: false
  warning: false
---

::: {.callout-tip}
## How labs work
You are encouraged to **work together and get started in class**---talk through the problems with the people around you. You will finish the lab on your own time and submit your own work.

**AI use: You may use AI to ask questions or debug, but not to write your code for you---the point is to think it through yourself. At the end of the lab, add a one-line note saying whether and how you used AI.**
:::

## Overview

In this lab you will make two kinds of maps: a static **choropleth** with `ggplot2` and `sf`, and an interactive **leaflet** map. We will map data relevant to Michael Ross's argument in [The Oil Curse](https://www-jstor-org.proxygw.wrlc.org/stable/j.ctt7s3wz.10) that oil wealth undermines democracy.

```{r}
#| label: setup

```

## Step 1: A choropleth map (50 pts)

*Use `rnaturalearth` to get country shapes (filter out Antarctica). Choose a variable relevant to Ross's argument---a democracy measure, oil rents (`NY.GDP.PETR.RT.ZS`), or taxes as a share of revenue---and join it to the shapes by ISO3C code. Then map it with `ggplot()` and `geom_sf()`, using `theme_map()` from `ggthemes` and a `viridis` color scale. Give the map clear labels. Below the map, say in a sentence or two what pattern you see.*

```{r}

```

## Step 2: A leaflet map of conflict (50 pts)

*Pick an oil-rich country that has experienced internal conflict (e.g., Iraq, Nigeria, Syria). Following the approach from the Interactive Maps topic, filter a few months of UCDP GED conflict data for your country and convert the coordinates to simple features with `st_as_sf()`. Then build a `leaflet` map with markers for the conflict events: show the location name on hover and the number of deaths and date on click. Use "OpenTopoMap" as your basemap.*

```{r}

```


<br>

**AI statement:** *(one line on whether and how you used AI)*
