---
title: "Lab 1: Plotting"
subtitle: "Do Wealth and Democracy Go Together?"
execute:
  echo: true
  message: false
  warning: false
---

::: {.callout-tip}
## How labs work
You are encouraged to **work together and get started in class**---talk through the problems with the people around you. You will finish the lab on your own time and submit your own work. You may use AI to **ask questions or debug**, but not to write your code for you; the point is to think it through yourself. At the end, add a one-line note saying whether and how you used AI.
:::

## Overview

In 1959, Seymour Martin Lipset published a [famous article](https://www-jstor-org.proxygw.wrlc.org/stable/1951731?seq=8) arguing that economic modernization---wealth, urbanization, education, health---makes democracy more likely. It has been cited more than 11,000 times. In this lab your group will put his argument to the test with three charts.

The data are already prepared for you, so you can focus entirely on the visualization. Run the chunk below to load them.

```{r}
#| label: setup

library(tidyverse)
library(scales)
library(plotly)
library(gganimate)

dem_data <- read_csv("data/dem_data.csv")

glimpse(dem_data)
```

Every row is one country in one year, from 1990 to 2023. Along with `country`, `year` and `region`, you have:

| Measures of democracy | Measures of modernization |
|----|----|
| `polyarchy` --- electoral democracy | `gdp_pc` --- GDP per capita (US\$) |
| `libdem` --- liberal democracy | `life_exp` --- life expectancy (years) |
| `partipdem` --- participatory democracy | `inf_mort` --- infant mortality (per 1,000) |
| `delibdem` --- deliberative democracy | `urban_pct` --- population that is urban (%) |
| `egaldem` --- egalitarian democracy | `internet_pct` --- population online (%) |

All five democracy measures run from 0 to 1, where higher is more democratic.

::: {.callout-note}
## Pick your variables
As a group, choose **one** democracy measure and **one** modernization measure to use for the whole lab. Different groups will choose differently, and that is the point---we will compare notes at the end.
:::

## Step 1: A line chart over time (30 pts)

*a) Pick three or four countries you are curious about, at different levels of wealth, and `filter()` the data down to just those countries.*

Country names have to match the data **exactly**. Run this to check a name before you use it---change `"Kor"` to the first few letters of whatever you are looking for:

```{r}
#| label: check-names

dem_data |>
  distinct(country) |>
  filter(str_detect(country, "Kor"))
```

::: {.callout-warning}
A few names are not what you would guess: `United States of America`, `Türkiye`, `Czechia`, `Burma/Myanmar`, and two separate Congos. Copy them from the check above rather than typing from memory.
:::

::: {.callout-note appearance="simple"}
## Starter code
Copy this into the chunk below and add your countries inside the `c()`.

```r
line_data <- dem_data |>
  filter(country %in% c("", ""))

glimpse(line_data)   # check you got what you expected
```
:::

```{r}
#| label: step1a

```

*b) Make a line chart. You need `year` on the x-axis, your democracy measure on the y-axis, and `country` mapped to `color` so each country gets its own line.*

::: {.callout-note appearance="simple"}
## Starter code
The `aes()` call is where you say which column maps to which part of the chart. Fill in all three, then add your labels inside `labs()`.

```r
ggplot(line_data, aes(x = , y = , color = )) +
  geom_line() +
  labs(x = , y = , title = , caption = )
```
:::

```{r}
#| label: step1b

```

*c) Now polish it. Add a theme and swap the default colors for a colorblind-safe scheme.*

::: {.callout-note appearance="simple"}
## Starter code
`country` is a discrete variable, so you want the discrete version of the viridis scale. Save the chart to an object first so you can keep building on it in the next part.

```r
line_chart <- ggplot(line_data, aes(x = , y = , color = )) +
  geom_line() +
  labs(x = , y = , title = , caption = ) +
  scale_color_viridis_d() +
  theme_minimal()

line_chart
```
:::

```{r}
#| label: step1c

```

*d) Mark a moment that matters. Add a dashed vertical line at a year when something important happened in one of your countries---an election, a coup, the start of a war---and label it with `annotate()`.*

::: {.callout-note appearance="simple"}
## Starter code
`xintercept` is the year the line crosses the x-axis. For the label, `x` and `y` are the coordinates where the text sits, so put it somewhere it will not collide with your lines.

```r
line_chart +
  geom_vline(xintercept = , linetype = "dashed") +
  annotate("text", x = , y = , label = "")
```
:::

```{r}
#| label: step1d

```

*e) In a sentence or two below this line, describe what you see. Which country changed the most? What does the event you marked in part (d) look like on the chart?*

## Step 2: A bar chart comparing regions (25 pts)

*a) Starting from `dem_data`, filter for a single recent year (2019 is a good choice), then group by `region` and take the mean of your democracy measure.*

::: {.callout-note appearance="simple"}
## Starter code
Three verbs in a row. Inside `summarize()` you give the new column a name, like `mean_polyarchy = mean(polyarchy)`.

```r
bar_data <- dem_data |>
  filter(year == ) |>
  group_by() |>
  summarize()

bar_data
```
:::

```{r}
#| label: step2a

```

*b) Make a bar chart with `geom_col()`, putting `region` on the x-axis and your summarized democracy measure on the y-axis. Add labels and a title. Add a color with `fill = ` and maintain the same theme as your first plot*

```{r}
#| label: step2b

```

*c) Below this line, say which region scores highest and which scores lowest.*

## Step 3: A scatter plot (25 pts)

*a) Filter `dem_data` for the same year you used in Step 2 and save it as `scatter_data`.*

```{r}
#| label: step3a

```

*b) Build a scatter plot: your modernization measure on the x-axis, your democracy measure on the y-axis, points colored by `region`. Then add a trend line with `geom_smooth(method = "lm")`, a colorblind-safe palette, a theme to match your other charts, and clear labels. Save it as `scatter_chart`---you will need it again in Step 4.*

Build it up one layer at a time---run the chart after each `+` you add so you can see what each piece does.

::: {.callout-warning}
## Where you put `color` matters here
Anything you put in the top-level `aes()` is inherited by **every** layer that follows it. So if you map `color = region` there, `geom_smooth()` will split by region too, and you will get six separate trend lines instead of one overall trend.

You want the points colored by region but a single trend line through all of them. Think about which layer actually needs the color, and put it there. If you end up with six trend lines, that is the clue.
:::

```{r}
#| label: step3b

```

::: {.callout-tip}
If you chose `gdp_pc`, most countries will bunch up against the left edge. Add this layer to spread them out:

`scale_x_log10(labels = label_dollar())`
:::

*c) Below this line, describe the relationship. Does it look positive, negative, or flat? If you used `inf_mort`, think about why its direction differs from the others.*

## Step 4: Make it interactive (10 pts)

*Take the scatter plot from Step 3 and make it interactive with `ggplotly()` from the `plotly` package. Then hover over a few points to see what the tooltip tells you.*

::: {.callout-note appearance="simple"}
## Starter code
By default the tooltip shows every variable you mapped. Use `tooltip =` to choose which ones appear, and add `aes(label = country)` if you want the country name in there---it is not in the plot otherwise.

```r
library(plotly)

ggplotly(scatter_chart)
```
:::

```{r}
#| label: step4

```

*Below this line, note one thing the interactive version lets you see that the static one did not.*

## Step 5: Conclusion (10 pts)

*Render your document, then write a short paragraph below this line. Taken together, do your charts support Lipset's argument? Name one country or region that does not fit the pattern.*

## Bonus (1 pt each)

*a) Add `facet_wrap(~ region)` to your scatter plot.*

*b) Animate your scatter plot across time with the `gganimate` package. Go back to the full `dem_data` rather than the single year you filtered to in Step 3, add `transition_states(year)`, and use `{closest_state}` inside your title so the year updates as it plays.*

*AI note: _(one line on whether/how you used AI)_*

::: {.callout-important}
## Submission Instructions

Head over to Blackboard and go to the Lab 1 assignment. Click "Create Submission" and write a brief statement saying that you have submitted the lab and that all of the work is your own.

Then upload a compressed (zipped) version of your project folder including the **rendered** HTML file. To compress your project folder, right-click it and choose Compress (Mac) or Send to → Compressed (zipped) folder (Windows).
:::
