---
title: "Lab 2: Wrangling"
execute:
  echo: true
  message: false
  warning: false
---

::: {.callout-tip}
## How labs work
You are encouraged to **work together and get started in class**---talk through the problems with the people around you. You will finish the lab on your own time and submit your own work.

**AI use: You may use AI to ask questions or debug, but not to write your code for you---the point is to think it through yourself. At the end of the lab, add a one-line note saying whether and how you used AI.**
:::

## Overview

In Lab 1 we handed you clean data. Now you will build a tidy data set yourself---the skill that makes every later visualization possible. You will pull a democracy measure from V-Dem and a development or empowerment indicator from the World Bank, wrangle them, merge them into one data frame, and finish with a chart comparing regions.

```{r}
#| label: setup

```

## Step 1: Download your data (20 pts)

*a) Use `vdem` from `vdemdata` to download the V-Dem dataset. Use `filter()` to keep the years from 1990 onward, then use `select()` to keep and rename the country name, the V-Dem country ID, the year, a democracy measure of your choice and the region variable (`e_regionpol_6C`). Recode the region numbers into names with `mutate()` and `case_match()`. `glimpse()` the result.*

```{r}
#| label: step1a

```

*b) Use `wb_data()` from `wbstats` to download GDP per capita (or another development or women's-empowerment indicator) for the same years. Give your indicator a readable name, drop `iso2c`, and rename `date` to `year` so it matches V-Dem. `glimpse()` the result.*

```{r}
#| label: step1b

```

## Step 2: Add country codes (15 pts)

*V-Dem and the World Bank use different country codes, so we need a common one before we can merge. Use `mutate()` and `countrycode()` to add an `iso3c` variable to your V-Dem data frame, converting from the `"vdem"` codes to the `"wb"` codes. `glimpse()` the result to confirm the new column is there.*

```{r}
#| label: step2

```

## Step 3: Merge (20 pts)

*Use `left_join()` to merge your V-Dem and World Bank data frames by `iso3c` and `year`. Make sure you end up with just one country name column. `glimpse()` the result and check for excessive `NA`s.*

```{r}
#| label: step3

```

## Step 4: Summarize by region (20 pts)

*a) Starting from your merged data frame, use `group_by()` and `summarize()` to calculate the mean (or median) of one of your variables---your democracy measure or your World Bank indicator---for each region. You can use all years or `filter()` for a single year first. Save the result as a new object.*

```{r}
#| label: step4a

```

*b) Use `arrange()` to sort the regions from highest to lowest. Which region comes out on top? Which is at the bottom?*

```{r}
#| label: step4b

```

## Step 5: Visualize (25 pts)

*a) Make a column chart with `geom_col()` showing your summarized variable by region. Sort the bars from highest to lowest, choose a fill color and a theme, and add axis labels, a title and a caption naming your source.*

```{r}
#| label: step5a

```

*b) Below this line, write a sentence or two describing what the chart shows. Is this what you expected?*

<br>

**AI statement:** *(one line on whether and how you used AI)*

---

## Hints

*Only look at these if you're stuck!*

**Hint 1 --- Recoding the region variable:**

`e_regionpol_6C` comes as numbers from 1 to 6. Use `case_match()` inside `mutate()` to turn them into names:

```r
mutate(
  region = case_match(region,
                      1 ~ "Eastern Europe",
                      2 ~ "Latin America",
                      3 ~ "Middle East",
                      4 ~ "Africa",
                      5 ~ "The West",
                      6 ~ "Asia")
)
```

**Hint 2 --- Naming your World Bank indicator:**

If you pass `wb_data()` a *named* vector, the column comes back with your name instead of the indicator code:

```r
indicators <- c("gdp_pc" = "NY.GDP.PCAP.CD")

wb_data(indicators, start_date = 1990, end_date = 2023)
```

**Hint 3 --- Do we need to reshape?**

No. Both V-Dem and `wb_data()` already give you one row per country-year, so the two data frames have the same shape. You only need to make sure the columns you join on have the same *names* (`year`, not `date`) and the same *codes* (that is what `countrycode()` is for).

**Hint 4 --- The `countrycode()` warning:**

You will probably see a warning like `Some values were not matched unambiguously`. That is OK. V-Dem includes a few historical units (like Zanzibar or Somaliland) that have no World Bank code, so they get an `NA` for `iso3c` and will not find a match in the merge.

**Hint 5 --- Two country columns after the merge:**

Both data frames have a `country` column, so `left_join()` gives you `country.x` and `country.y`. Drop the World Bank version with `select(-country)` before you join, or drop and rename the extra column after.

**Hint 6 --- `NA`s in your summary:**

If your summary comes back as `NA`, it is because some country-years are missing data. Add `na.rm = TRUE`:

```r
summarize(gdp_pc = mean(gdp_pc, na.rm = TRUE))
```

**Hint 7 --- Sorting the bars:**

`arrange()` sorts your table but not your chart---`ggplot()` will still put the regions in alphabetical order. Use `fct_reorder()` inside `aes()` to sort the bars (the `-` puts the highest value first):

```r
ggplot(region_summary, aes(x = fct_reorder(region, -gdp_pc), y = gdp_pc)) +
  geom_col(fill = "steelblue")
```

**Hint 8 --- Formatting the y-axis:**

If you chart GDP per capita, `scale_y_continuous(labels = label_dollar())` from the `scales` package will format the axis as dollars. For a percentage, use `label_percent(scale = 1)`.
